---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  warning = FALSE,
  message = FALSE,
  dev = "ragg_png",
  dpi = 300)
```

# z22

<!-- badges: start -->
[![R-CMD-check](https://github.com/JsLth/z22/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/JsLth/z22/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/z22)](https://CRAN.R-project.org/package=z22)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Codecov test coverage](https://codecov.io/gh/JsLth/z22/branch/main/graph/badge.svg)](https://app.codecov.io/gh/JsLth/z22?branch=main)
[![CodeFactor](https://www.codefactor.io/repository/github/jslth/z22/badge/main)](https://www.codefactor.io/repository/github/jslth/z22/overview/main)
<!-- badges: end -->

As part of the German census, a number of socio-economic indicators are
published at a small-scale grid level (100m, 1km, 10km). Usually the
corresponding data files are very large and hard to work with. `{z22}` aims to
quickly and efficiently retrieve German census grid data. The package also
eases comparisons to grid data from the census 2011 by establishing a
standardized feature format.

Powered by the [z22data](https://github.com/jslth/z22data) data repository
which hosts the census grid data in small and digestible chunks. Based on the
predecessor package [z11](https://github.com/stefanjuenger/z11/) which allows
easy access to the gridded data from the Census 2011.


## Installation

You can install the development version of z22 like so:

```{r eval=FALSE}
pak::pkg_install("jslth/z22")
```


## Example

The package contains a main function, `z22_data`, to retrieve and process census
grid data. The following example downloads a feature grid on the vacancy share
in Germany at a 1x1 km resolution.

```{r}
library(z22)
library(terra)

grid_vac <- z22_data("vacancies", res = "10km", rasterize = TRUE)
grid_vac
```

<details><summary>Code for the plot</summary>
```{r eval=FALSE}
library(ggplot2)

df_vac <- z22_pivot_longer(grid_vac, "vacancies")
ggplot(df_vac) +
  geom_tile(aes(x, y, fill = value)) +
  coord_sf(crs = 3035) + 
  facet_wrap(~category, nrow = 2) +
  scale_fill_viridis_c("Share", na.value = "transparent", transform = "log2") +
  theme_bw() +
  labs(x = NULL, y = NULL) +
  theme(panel.grid = element_blank(), axis.text = element_blank())
```
</details>

```{r echo=FALSE}
library(ggplot2)

df_vac <- z22_pivot_longer(grid_vac, "vacancies")
ggplot(df_vac) +
  geom_tile(aes(x, y, fill = value)) +
  coord_sf(crs = 3035) + 
  facet_wrap(~category, nrow = 2) +
  scale_fill_viridis_c("Count", na.value = "transparent", transform = "log2") +
  theme_bw() +
  labs(x = NULL, y = NULL) +
  theme(panel.grid = element_blank(), axis.text = element_blank())
```

## Available data

```{r echo=FALSE}
knitr::kable(z22:::make_feature_table())
```

